---
title: "Classification - Neo Zhao - CS4375"
output:
  pdf_document: default
---

## Linear Models
* Logistic Regression uses a qualitative target variable to predict. In this project, I have found the ratings of Red and White wine. I will be setting all ratings > 3 to 1 and all ratings < 3 to 0. While there were about 32 observations that were exactly 3, we will omit them as it will not mess with the data too much out of 12,000+ observations. The Linear Model for classification will create a sort of barrier to seperate into different classes. In this project, Ratings > 3 and Ratings < 3 will be predicted into 2 different classes. 

```{r}
library(tidyverse)
library(dplyr)
library(ROCR)
library(mccr)
library(ISLR)
library(caret)
library(tree)
library(rpart)

# Source: https://www.kaggle.com/datasets/budnyak/wine-rating-and-price?select=Red.csv 

# Red, White, Rose, and Sparkling wine are all from the same dataset; however, separated by type

# Red Total: 8666 
Red <- read.csv("Red.csv")

# White Total: 3764
White <- read.csv("White.csv")

# Rose Total: 397
Rose <- read.csv("Rose.csv")

# Sparkling Total: 1007
Sparkling <- read.csv("Sparkling.csv")

# Combine the datasets together, Total: 13058
totalWine <- rbind(data = Red, data = White, data = Rose, data = Sparkling)

# Rename ï..Name to just Name
names(totalWine)[1] <- "Name"

# Omit Names, Winery, & Region Column
totalWine <- subset(totalWine, select = -c(Name, Winery, Region))

# Omit all records where Year = N.V.
totalWine <- subset(totalWine, totalWine$Year != "N.V.")

# Omit all records where Rating = 3, Total: 12398
totalWine <- subset(totalWine, totalWine$Rating != 3)

# Omit all records before 2000s
totalWine <- subset(totalWine, totalWine$Year >= 2000)

# Replace ratings with 1 if Rating > 3 and replace with 0 if Rating < 3
totalWine$Rating[totalWine$Rating <= 3] <- 0
totalWine$Rating[totalWine$Rating > 3] <- 1

# Reorder Columns
totalWine <- totalWine[,c(1,2,3,5,4)]
```

```{r}
# Split to 80/20 Train/Test
set.seed(512)
i <- sample(1 : nrow(totalWine), round(nrow(totalWine) * 0.8), replace = FALSE)
train <- totalWine[i,]
test <- totalWine[-i,]
```

### Data Exploration
```{r}
# 1) summary()
summary(train)
```

```{r}
# 2) is.na()
colSums(is.na(train))
```

```{r}
# 3) str()
str(train)
```

```{r}
# 4) head() functions
head(train)
```

```{r}
# 5) cor() and pairs()
cor(train[,c(-1, -4)])

pairs(train[,c(-1, -4)])
```

### Informative Graphs
```{r}
# Red
plot(Rating ~ Price, data = Red, main = "Red Wine", xlab = "Price", ylab = "Rating")

# White
plot(Rating ~ Price, data = White, main = "White Wine", xlab = "Price", ylab = "Rating")

# Rose
plot(Rating ~ Price, data = Rose, main = "Rose Wine", xlab = "Price", ylab = "Rating")

# Sparkling
plot(Rating ~ Price, data = Sparkling, main = "Sparkling Wine", xlab = "Price", ylab = "Rating")
```

### Linear Regression
* The correlation for the Linear Model was 28%, which is not the best. Can kNN do better?
```{r}
# Split to 80/20 Train/Test
set.seed(512)
i <- sample(1 : nrow(totalWine), round(nrow(totalWine) * 0.8), replace = FALSE)
train <- totalWine[i,]
test <- totalWine[-i,]
```

```{r}
# Linear Regression
lm1 <- lm(Price ~ ., data = train)
summary(lm1)
```

```{r}
# Plotting Residuals
par(mfrow = c(2,2))
plot(lm1)
```

```{r}
pred1 <- predict(lm1, newdata = test)
cor1 <- cor(pred1, test$Price)
mse1 <- mean((pred1 - test$Price) ^ 2)
rmse1 <- sqrt(mse1)

print(paste ("Cor = ", cor1))
print(paste ("MSE = ", mse1))
```

### kNN Clustering - Regression
* The correlation for the knn Regression was 18%, which is lower than Linear Regression. The MSE is slightly higher than the Linear Regression; however, they are both very very high. Let's scale the data to see if it does any better.
* The scaled correlation is 16%, which is only slightly higher and does not compete with the Linear Regression correlation. The MSE also is higher by 200.
```{r}
# Linear Model with all Predictors
lm1 <- lm(Price ~. , data = totalWine)
summary(lm1)
```


```{r}
# kNN Regression
train$Year <- as.numeric(train$Year)
test$Year <- as.numeric(test$Year)

fit <- knnreg(train[,2:4], train[,5], k = 3)
predictions <- predict(fit, test[,2:4])
cor_knn1 <- cor(predictions, test$Price)
mse_knn1 <- mean((predictions - test$Price) ^ 2)

print(paste ("Cor = ", cor_knn1))
print(paste ("MSE = ", mse_knn1))
```

```{r}
# Scaled Data
trainScaled <- train[,2:4]
means <- sapply(trainScaled, mean)
stdvs <- sapply(trainScaled, sd)
trainScaled <- scale(trainScaled, center = means, scale = stdvs)
testScaled <- scale(test[,2:4], center = means, scale = stdvs)

fit <- knnreg(trainScaled, train$Price, k = 3)
predictions <- predict(fit, testScaled)
cor_knn2 <- cor(predictions, test$Price)
mse_knn2 <- mean((predictions - test$Price) ^ 2)

print(paste ("Cor = ", cor_knn2))
print(paste ("MSE = ", mse_knn2))
```

```{r}
# Find the best k
cor_k <- rep(0, 20)
mse_k <- rep(0, 20)
i <- 1
for(k in seq(1, 39, 2)) {
  fit_k <- knnreg(trainScaled, train$Price, k = k)
  pred_k <- predict(fit_k, testScaled)
  cor_k[i] <- cor(pred_k, test$Price)
  mse_k[i] <- mean((pred_k - test$Price) ^ 2)
  print(paste("k =", k, cor_k[i], mse_k[i]))
  i <- i + 1
  }
```

```{r}
plot(1:20, cor_k, lwd = 2, col = 'red', ylab = "", yaxt = 'n')
par(new = TRUE)
plot(1:20, mse_k, lwd = 2, col = 'blue', labels = FALSE, ylab = "", yaxt = 'n')
```

```{r}
# Let's Pick a different k = 39

fit_39 <- knnreg(trainScaled, train$Price, k = 39)
predictions_39 <- predict(fit_39, testScaled)
cor_knn39 <- cor(predictions_39, test$Price)
mse_knn39 <- mean((predictions_39 - test$Price) ^ 2)

print(paste ("Cor = ", cor_knn39))
print(paste ("MSE = ", mse_knn39))
```

### Decision Trees
* My last code block makes another tree. The accuracy is an insane 99.8%!
```{r}
# Split to 80/20 Train/Test again just for fun I guess
set.seed(1002)
i <- sample(1 : nrow(totalWine), round(nrow(totalWine) * 0.8), replace = FALSE)
train <- totalWine[i,]
test <- totalWine[-i,]
```

```{r}
# Using Tree
tree1 <- tree(Price ~ ., data = train, method = "class")
summary(tree1)
```
```{r}
# Plotting Tree pt. 1
plot(tree1, uniform = TRUE, margin = 0.2)
text(tree1)
```

```{r}
# Hehe Tree Pt. 2
tree2 <- tree(Rating ~ ., data = train)
tree2

summary(tree2)
```

```{r}
# Plotting Tree pt. 2
plot(tree2)
text(tree2, cex = 0.75, pretty = 0)
```

```{r}
set.seed(4375)
i <- sample(150, 100, replace = FALSE)
train <- totalWine[i,]
test <- totalWine[-i,]
tree3 <- tree(Rating ~ ., data = train)
pred <- predict(tree3, newdata = test)
table(pred, test$Rating)
```

```{r}
mean(pred == test$Rating)
```





